Forecasting the old-age dependency ratio to determine the best pension age

### 2020-05-05 PensionAge.tex
Now compiling

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}%
        \vspace*{-2cm}
        \centerline{\raisebox{-1.8cm}{\includegraphics[width=5cm]{MBSportrait}}\hspace*{9cm} ISSN 1440-771X}\vspace{0.99cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \scriptsize http://business.monash.edu/econometrics-and-business-statistics/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

@@

@@

\begin{document}
\maketitle
\begin{abstract}
We forecast the old-age dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old-age dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. We apply these ideas to Australian historical demographic data and the results suggest that the pension age should be increased to 67 by 2021, 70 by 2032 and 75 by 2058 in order to maintain the old-age dependency ratio at the current level, which is 21\%. This approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries.
\end{abstract}
\begin{keywords}
functional time series model, coherent forecasts, demographic components, pension and insurance
\end{keywords}

@@

This is consistent with the Australian Intergenerational Reports \autocites{IGR02}{IGR07}{Swan2010} that explored the extent to which existing age-related policies would affect the burden of future working generations. To date, the Australian government has addressed the problem through a variety of approaches including modifications to compulsory superannuation policies, instigating work incentives for older people, and increasing the pension age (defined as the age at which eligible residents can access the age pension for the first time). In 2009, the federal government announced an increase in pension age from 65 to 67 by 2023 \autocite{Nielson2010}. In May 2014, the Federal Treasurer proposed a further increase in the pension age to 70 by 2035. This new proposal is in line with the recommendations from the \textcite{PC13} and the \textcite{NCOA2014}. In this paper, we investigate the effect of these changes in the pension age on the forecast OADR.

@@

In this paper, we use the OADR as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs \autocites{ARW07}{IGR02}{IGR07}. The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age \autocite{ARW07}. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method \autocite{HU07} and the product-ratio method \autocite{HBY13} to forecast a population age structure. Forecasts of population structure have been built on forecasts of age- and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age- and gender-specific population structure.

@@

\hypertarget{sec:litreview}{%
\section{Background and Literature Review}\label{sec:litreview}}

@@

Age pension in Australia is an income support and provides access to a range of concessions for older Australians who have passed means testing. For example, it provides single home-owners with a real income of \$13653 p.a and additional allowances to cover expenses such as rent or pharmaceuticals\footnote{The pension amount is adjusted every 6 months for inflation}. The means test ensures that retirees with enough personal savings will become ineligible for public provision by reducing the pension at fixed levels of pay as personal wealth increases. The age pension in Australia is very generous, even with the means test. For instance, a couple holding assets of more than one million dollars is still eligible for a partial age pension.

Currently, the age pension is the primary source of income for 70\% -80\% of Australian residents over 64 years \autocite{ABS06}. Even though the government has promoted self-financed retirement by introducing the Superannuation Guarantee in 1992, the Superannuation Guarantee accumulations are still modest, with the current average less than \$100,000 at retirement \autocite{ASFA07}. It has been estimated by \textcite{PT08} that a retiree will need to have \$450,000 at 65 in order to support a pension-equivalent income and this is more than eight times of the average retirement accumulation of Australians. The required wealth to comply with personal-funded retirement savings requirements is too high for most Australians, suggesting that retirees will continue to rely on the age pension in the coming decades.

@@

The issue is not only a problem for the government, but also for future retirees. Given children born today are expected to live up to their 90s on average, the \textcite{PC13} suggests that Australians won't be able to afford to spend 35 years in retirement (i.e., retire around 65). They have predicted that a third of baby boomers are expected to spend all their money and assets before they die\footnote{Baby Boomers are people born during the Post-World War II baby boom between the years 1946 and 1964}. This problem might eventually deteriorate the sustainability of the age pension, although the welfare system works well at the moment, because demographic changes often have delayed effects \autocite{IGR02}.

@@

The Australian government is trying to take early intervention in relation to the pension age to balance between providing social support and fiscal cost. The pension age for males has been set at 65 for over a century. While the pension age for females was 60 years until 1994, it has steadily increased to 65 years, in line with that of males. The previous Labor government announced in 2009 that the pension age will increase six months every two years until 2017, capping at 67 in July 2023. This change is expected to have a significant impact on labor force participation and to decrease the retirement duration of residents because pension age is regarded as a normal age for retirement in Australia. However, the \textcite{PC13} proposed further lifting of the pension age to 70 on the grounds that increasing the pension age to 70 would save taxpayers \$150 billion in welfare costs and health spending, and elderly people are capable of continuing work. The National Commission of Audit (NCOA) (2014) holds the same opinion, and recommended lifting the pension age to 70 from 2053. The current Treasurer Joe Hockey announced that the government would increase the pension age to 70 from 2035, which is proposed in the Federal Budget Statement for 2014-2015 to be confirmed by Australian Federal Government. Figure \ref{fig:changespension} illustrates the process of the adjustments on pension age scheme. This announcement received overwhelmingly negative responses in the media. \textcite{Guest14} argues from an economic prospective that Australia currently has much better circumstances than other OECD countries in terms of age pension pressure, because spending on the age pension in Australia is less than half of the OECD average. He also claims that Australia has a higher consumption of all other goods and services, and so why are they not allowed to have more leisure time. Further, \textcite{Power14} contends that many Australians are not physically able to work full-time to 70, so that recommendations made by the PC and NCOA are simplistic solutions to a more a complex social problem.

@@

\hypertarget{deterministic-methods-used-by-the-abs}{%
\subsubsection{Deterministic Methods Used by the ABS}\label{deterministic-methods-used-by-the-abs}}

The deterministic method is judgment-based. The Australian Bureau of Statistics formulates assumptions on the basis of demographic trends, in conjunction with advice from individual experts or government departments. The projections are not predictions or forecasts, ``but are simply illustrations of the growth and change in population which would occur if certain assumptions about future levels of fertility, mortality, internal migration and overseas migration were to prevail over the projection period'' \autocite[p.2]{ABS12}. In addition, they do not attempt to take into account of non-demographic factors, such as major government policy changes and catastrophes. Therefore, they normally report multiple (usually three) series based on different levels of assumptions of demographic factors to allow for variety of population projections.

\hypertarget{lee-carter-method}{%
\subsubsection{Lee-Carter Method}\label{lee-carter-method}}

@@

The model proposed by \textcite{LC92} is as follows:
\begin{equation}
  \ln(m_{x,t})=a_x+b_xk_t+\epsilon_{x,t}
\end{equation}
where \(m_{x,t}\) is the central death rate\footnote{The central death rate is properly defined in Section~\ref{sec:data}} for age \(x\) in year \(t\). This model involves fitting a matrix of log central death rates with sets of age-specific constants, \{\(a_x\)\} and \{\(b_x\)\}, and a time-varying index \(k_t\). \(b_x\) is a set of parameters that tell us the rate at which each age responds to changes in \(k_t\). \(k_t\) measures the general level of the log death rates; \(a_x\) is the general pattern across age of the log mortality rate and \(\epsilon_{x,t}\) is an error term with zero mean and constant variance, reflecting the randomness that can't be captured by the model.

The model is under-determined in terms of estimation of parameters. For example, if the vector \(a,b,k\) is one solution, then for any scalar \(c\), \(a, b/c, kc\) or \(a - bc, b, k+c\) also must be two possible solutions. Therefore, Lee and Carter (1992) impose the following constraints:
\begin{equation}
  \sum_{t=1}^{n}k_t=0, \qquad \sum_{x=x_1}^{x_m}b_x=1
\end{equation}
Also, we normally seek the least squares solution when estimating parameters. However, in this case, the model cannot be estimated using simple regression methods because there is no regressor on the right side of the equation. \textcite{LC92} choose a singular value decomposition (SVD) method to undertake principal component analysis (PCA), and hence find the least squares solution. It has been shown that applying SVD method to a matrix of the logarithms of the rates after subtracting the average over time of the log age-specific rates can find the least square solution (Good, 1969). Therefore, the estimations of \(b_x\) and \(k_t\) are the first principal component and scores of \(\ln(m_{x,t})-a_x\) (i.e., \(b_x=\bm{z_1}\) and \(k_t=\bm{u_1}'\) when \(\bm{X}=\ln(m_{x,t})-a_x\)). However, the empirical results suggest that the resulting fitted number of deaths are biased estimates of actual number of deaths because of the \(\ln\) transformation. To overcome this problem, the LC method adjusts \(k_t\) such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted \(k_t\) is then extrapolated using random walk with drift models, which is equivalent to ARIMA (\(0,1,0\)) model with \(c\neq0\).

\hypertarget{hyndman-ullah-method}{%
\subsubsection{\texorpdfstring{Hyndman-Ullah Method \label{sec:HU}}{Hyndman-Ullah Method }}\label{hyndman-ullah-method}}

@@

\begin{enumerate}
\item The rates are smoothed using penalized regression splines [@BHT14] before modeling.
\item It uses functional principal components analysis, which is a continuous version of PCA.
\item It uses more than one principal component, which address the main weakness of the LC method.
\item The forecasting models for the principal component scores are more complex than the Random walk with drift model used in LC model.
\end{enumerate}

The model proposed by \textcite{HU07} is as follows:
\vspace{-.3in}
\begin{align}
  y_t(x) & = s_t(x) + \sigma_t(x)\epsilon_{t,x}\label{eq:HU_1}\\
  s_t(x) & = \mu(x) + \sum_{j=1}^{J}\beta_{t,j}\phi_j(x) + e_t(x)\label{eq:HU_2}
\end{align}
Equation \eqref{eq:HU_1} describes the smoothing process. \(y_t(x)\) is the observed data in year \(t\) and age \(x\), and \(s_t(x)\) is the smoothed function of \(y_t(x)\), estimated by constrained weighted penalized regression splines \autocite{BHT14}. \(\sigma_t(x)\) allows the variance to be a function of time and age; and \(\epsilon_{t,x}\) are independent identical distributed standard normal error terms. Equation \eqref{eq:HU_2} has a similar structure to the LC method, except that it incorporates more than one principal component and score. It describes the dynamics of \(s_t(x)\) over time. The \(\beta_{t,j}\) for every \(j\) and \(\mu(x)\) are sets of age-specific constants, and \(\phi_j(x)\) for every \(j\) is a time-varying index. The \(e_t(x)\) is the model error, which is the residual function with mean zero and no serial correlation. The \(\mu(x)\) is estimated by the average of \(s_t(x)\) across years, which measures the general shape of the observed data \(y_t(x)\). The estimates of \(\phi_j(x)\) and \(\beta_{t,j}\) are the \(j\)th principal component and scores of \(s_t(x)-\mu(x)\) respectively. In other words, we take singular value decomposition of \(s_t(x)-\mu(x)\), then \eqref{eq:HU_2} now becomes
\vspace{-.2in}
\begin{equation*}
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_J(x)\beta_{t,J}
\end{equation*}
We chose \(J\) in a way that we believe the majority of information for \(s_t(x)-\mu(x)\) has been captured by the first \(J\) components.

In terms of forecasting, by conditioning on the observed data \(\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}\) and the set of functional principal components \(\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}\), the \(h\)-step-ahead forecast of log rate \(y_{n+h}(x)\) can be acquired by:
\vspace{-.2in}
\begin{equation}
  \widehat{y}_{n+h\mid n}(x)=\widehat{s}_{n+h\mid n}(x) = \text{E}[\widehat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \widehat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\widehat{\beta}_{n+h\mid n,j}
\end{equation}
where \(\widehat{\beta}_{n+h\mid n,j}\) denotes the \(h\)-step-ahead forecast of \(\beta_{n+h\mid n,j}\). Because of the time-varying property of \(\left\{\beta_{t,j}\right\}\), it controls the dynamics of the process. For every component series \(\left\{\beta_{t,j}\right\}\), we fit them with an ARIMA\((p_j,d_j,q_j)\) process, where \(p_j,d_j,q_j\) are determined by minimizing the Akaike Information Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process is given as follows:
\vspace{-.3in}
\begin{equation*}
  (1-\phi_1B-\dots-\phi_{p_j}B^{p_j})(1-B)^{d_j}\beta_{t,j}=c+(1+\theta_1B+\dots+\theta_{q_j}B^{q_j})e_t, \quad ~\forall j=1,2,\dots,J,
\end{equation*}
where the parameters are estimated using maximum likelihood estimation.

@@

We use the same notation and similar procedures to construct data as in \textcite{HB08}, because the fundamental model for forecasting population for this study is the \textcite{HU07} method. The data used to construct population age structure included: age- and gender-specific birth and death numbers of each calendar year, age- and gender-specific population members on 1~January of each year, and age- and gender-specific exposures to risk (i.e., population of age \(x\) at 30 June) for each year. The notations are as follows:
\vspace{-.4in}
\begin{align*}
  B_t(x) &= \text{Births in calendar year $t$ to females of age $x$}\\
  D_t(x) &= \text{Deaths in calendar year $t$ of persons of age $x$}\\
  P_t(x) &= \text{Population of age $x$ at 1 January of year $t$}\\
  E_t(x) &= \text{Population of age $x$ exposed to risk at 30 June of year $t$}\\
\end{align*}
\vspace{-.4in}where \(x\) denotes age and \(x=0,1,2,\dots,m^+\). \(m^+\) is the open-ended upper age group, which is 100 in this study. \(t\) denotes year and \(t=1,2,\dots,n\), where \(n\) is the horizon of historical data used. Therefore, all of those data points can be recorded in an \(m\times n\) matrix. The superscripts M and F denote male and female respectively.

We can obtain historical mortality rates \(m_t(x)\) (also called central death rates) and fertility rates \(f_t(x)\) as follows:
\vspace{-.3in}
\begin{align*}
  m_t(x) & = \frac{D_t(x)}{E_t(x)}=\text{age-sex-specific central death rates in calendar year t}\\
  f_t(x) & = \frac{B_t(x)}{E^F_t(x)}=\text{age-specific fertility rates in calendar year t}
\end{align*}

>>>>

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \placefig{2}{1.5}{width=5cm}{monash2}
        \placefig{16.9}{1.5}{width=2.1cm}{MBusSchool}
        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
        \begin{textblock}{7}(12.7,27.9)\hfill
        \includegraphics[height=0.7cm]{AACSB}~~~
        \includegraphics[height=0.7cm]{EQUIS}~~~
        \includegraphics[height=0.7cm]{AMBA}
        \end{textblock}
        \vspace*{2cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \footnotesize http://monash.edu/business/ebs/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

@@

\RequirePackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}
\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}

@@

\begin{document}
\maketitle
\begin{abstract}
We forecast the old-age dependency ratio for Australia under various pension age proposals, and estimate a pension age scheme that will provide a stable old-age dependency ratio at a specified level. Our approach involves a stochastic population forecasting method based on coherent functional data models for mortality, fertility and net migration, which we use to simulate future age-structures of the population. We apply these ideas to Australian historical demographic data and the results suggest that the pension age should be increased to 67 by 2021, 70 by 2032 and 75 by 2058 in order to maintain the old-age dependency ratio at the 2014 level, which is 21\%. This approach can easily be extended to other target levels of the old-aged dependency ratio and to other countries.
\end{abstract}
\begin{keywords}
functional time series model, coherent forecasts, demographic components, pension and insurance
\end{keywords}

@@

This is consistent with the Australian Intergenerational Reports \autocite{IGR02,IGR07,Swan2010} that explored the extent to which existing age-related policies would affect the burden of future working generations. To date, the Australian government has addressed the problem through a variety of approaches including modifications to compulsory superannuation policies, instigating work incentives for older people, and increasing the pension age (defined as the age at which eligible residents can access the age pension for the first time). In 2009, the federal government announced an increase in pension age from 65 to 67 by 2023 \autocite{Nielson2010}. In May 2014, the Federal Treasurer proposed a further increase in the pension age to 70 by 2035. This new proposal is in line with the recommendations from the \textcite{PC13} and the \textcite{NCOA2014}. In this paper, we investigate the effect of these changes in the pension age on the forecast OADR.

@@

In this paper, we use the OADR as a measure of the financial burden on the Australian workforce because a rise in the OADR will reduce the ability of working age population to finance pensions and health costs \autocite{ARW07,IGR02,IGR07}. The OADR is defined as the ratio of the number of people aged over the pension age to the number of people of working age \autocite{ARW07}. This can be expressed using the formula below:
\begin{equation}
  \text{OADR} = \frac{\text{number of people aged over pension age}}{\text{number of people aged 15 to pension age}}*100\%
\end{equation}
We provide forecasts of future population age structures over the next 50 years. We then use those forecasts to compute the OADR associated with pension age both before and after the adjustments. One contribution of the current study is the combined use of the Hyndman-Ullah forecasting method \autocite{HU07} and the product-ratio method \autocite{HBY13} to forecast a population age structure. Forecasts of population structure have been built on forecasts of age- and gender-specific mortality rates, fertility rates and net migration using this method. These forecasts presuppose that the historical trends of past years will continue. Historical data of those rates have thus been used as input to simulate future paths of age- and gender-specific population structure.

@@

\hypertarget{sec:litreview}{%
\section{Background}\label{sec:litreview}}

@@

Age pension in Australia is an income support and provides access to a range of concessions for older Australians who have passed means testing. For example, it provides single home-owners with a real income of \$23,889 p.a and additional allowances to cover expenses such as rent or pharmaceuticals\footnote{The pension amount is adjusted every 6 months for inflation}. The means test ensures that retirees with enough personal savings will become ineligible for public provision by reducing the pension at fixed levels of pay as personal wealth increases. The age pension in Australia is very generous, even with the means test. For instance, a couple holding assets of more than one million dollars is still eligible for a partial age pension.

In 2006, the age pension was the primary source of income for 70--80\% of Australian residents over 64 years \autocite{ABS06}. Even though the government has promoted self-financed retirement by introducing the Superannuation Guarantee in 1992, the Superannuation Guarantee accumulations are still modest, with the average less than \$100,000 at retirement \autocite{ASFA07}. It has been estimated by \textcite{PT08} that a retiree will need to have \$450,000 at 65 in order to support a pension-equivalent income and this is more than eight times of the average retirement accumulation of Australians. The required wealth to comply with personal-funded retirement savings requirements is too high for most Australians, suggesting that retirees will continue to rely on the age pension in the coming decades.

@@

The issue is not only a problem for the government, but also for future retirees. Given children born today are expected to live up to their 90s on average, the \textcite{PC13} suggests that Australians won't be able to afford to spend 35 years in retirement (i.e., retire around 65). They have predicted that a third of baby boomers are expected to spend all their money and assets before they die\footnote{Baby Boomers are people born during the Post-World War II baby boom between the years 1946 and 1964.}. This problem might eventually deteriorate the sustainability of the age pension, although the welfare system works well at the moment, because demographic changes often have delayed effects \autocite{IGR02}.

@@

The Australian government is trying to take early intervention in relation to the pension age to balance between providing social support and fiscal cost. The pension age for males has been set at 65 for over a century. While the pension age for females was 60 years until 1994, it has steadily increased to 65 years, in line with that of males. The previous Labor government announced in 2009 that the pension age will increase six months every two years until 2017, capping at 67 in July 2023. This change is expected to have a significant impact on labor force participation and to decrease the retirement duration of residents because pension age is regarded as a normal age for retirement in Australia. However, the \textcite{PC13} proposed further lifting of the pension age to 70 on the grounds that increasing the pension age to 70 would save taxpayers \$150 billion in welfare costs and health spending, and elderly people are capable of continuing work. The National Commission of Audit (NCOA) (2014) holds the same opinion, and recommended lifting the pension age to 70 from 2053. The former Treasurer Joe Hockey announced that the government would increase the pension age to 70 from 2035, which is proposed in the Federal Budget Statement for 2014--2015 to be confirmed by Australian Federal Government. Figure \ref{fig:changespension} illustrates the process of the adjustments on pension age scheme. This announcement received overwhelmingly negative responses in the media. \textcite{Guest14} argues from an economic prospective that Australia currently has much better circumstances than other OECD countries in terms of age pension pressure, because spending on the age pension in Australia is less than half of the OECD average. He also claims that Australia has a higher consumption of all other goods and services, and so why are they not allowed to have more leisure time. Further, \textcite{Power14} contends that many Australians are not physically able to work full-time to 70, so that recommendations made by the PC and NCOA are simplistic solutions to a more a complex social problem.

@@

\hypertarget{deterministic-methods-used-by-the-abs}{%
\subsubsection{Deterministic methods used by the ABS}\label{deterministic-methods-used-by-the-abs}}

The deterministic method is judgment-based. The Australian Bureau of Statistics formulates assumptions on the basis of demographic trends, in conjunction with advice from individual experts or government departments. The projections are not predictions or forecasts, ``but are simply illustrations of the growth and change in population which would occur if certain assumptions about future levels of fertility, mortality, internal migration and overseas migration were to prevail over the projection period'' \autocite[p.2]{ABS12}. In addition, they do not attempt to take into account of non-demographic factors, such as major government policy changes and catastrophes. Therefore, they normally report multiple (usually three) series based on different levels of assumptions of demographic factors to allow for a variety of population projections.

\hypertarget{lee-carter-method}{%
\subsubsection{Lee-Carter method}\label{lee-carter-method}}

@@

The model proposed by \textcite{LC92} is as follows:
\begin{equation}
  \log(m_{x,t})=a_x+b_xk_t+\epsilon_{x,t}
\end{equation}
where \(m_{x,t}\) is the central death rate (properly defined in Section\textasciitilde\ref{sec:data}) for age \(x\) in year \(t\). This model involves fitting a matrix of log central death rates with sets of age-specific constants, \{\(a_x\)\} and \{\(b_x\)\}, and a time-varying index \(k_t\). \(b_x\) is a set of parameters that tell us the rate at which each age responds to changes in \(k_t\). \(k_t\) measures the general level of the log death rates; \(a_x\) is the general pattern across age of the log mortality rate and \(\epsilon_{x,t}\) is an error term with zero mean and constant variance, reflecting the randomness that can't be captured by the model.

The model is under-determined in terms of estimation of parameters. For example, if the vector \(a,b,k\) is one solution, then for any scalar \(c\), \(a, b/c, kc\) or \(a - bc, b, k+c\) also must be two possible solutions. Therefore, Lee and Carter (1992) impose the following constraints:
\begin{equation}
  \sum_{t=1}^{n}k_t=0, \qquad \sum_{x=x_1}^{x_m}b_x=1
\end{equation}
Also, we normally seek the least squares solution when estimating parameters. However, in this case, the model cannot be estimated using simple regression methods because there is no regressor on the right side of the equation. \textcite{LC92} choose a singular value decomposition (SVD) method to undertake principal component analysis (PCA), and hence find the least squares solution. It has been shown that applying SVD method to a matrix of the logarithms of the rates after subtracting the average over time of the log age-specific rates can find the least squares solution (Good, 1969). Therefore, the estimates of \(b_x\) and \(k_t\) are the first principal component and scores of \(\log(m_{x,t})-a_x\) (i.e., \(b_x=\bm{z_1}\) and \(k_t=\bm{u_1}'\) when \(\bm{X}=\log(m_{x,t})-a_x\)). However, the empirical results suggest that the resulting fitted number of deaths are biased estimates of actual number of deaths because of the \(\log\) transformation. To overcome this problem, the LC method adjusts \(k_t\) such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted \(k_t\) is then extrapolated using random walk with drift models, which is equivalent to an ARIMA (\(0,1,0\)) model with \(c\neq0\).

\hypertarget{hyndman-ullah-method}{%
\subsubsection{\texorpdfstring{Hyndman-Ullah method \label{sec:HU}}{Hyndman-Ullah method }}\label{hyndman-ullah-method}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The rates are smoothed using penalized regression splines \autocite{BHT14} before modeling.
\item
  It uses functional principal components analysis, which is a continuous version of PCA.
\item
  It uses more than one principal component, which address the main weakness of the LC method.
\item
  The forecasting models for the principal component scores are more complex than the Random walk with drift model used in LC model.
\end{enumerate}

The model proposed by \textcite{HU07} is as follows:
\begin{align}
  y_t(x) & = s_t(x) + \sigma_t(x)\epsilon_{t,x}\label{eq:HU_1}\\
  s_t(x) & = \mu(x) + \sum_{j=1}^{J}\beta_{t,j}\phi_j(x) + e_t(x)\label{eq:HU_2}
\end{align}
Equation \eqref{eq:HU_1} describes the smoothing process. \(y_t(x)\) is the observed data in year \(t\) and age \(x\), and \(s_t(x)\) is the smoothed function of \(y_t(x)\), estimated by constrained weighted penalized regression splines \autocite{BHT14}. \(\sigma_t(x)\) allows the variance to be a function of time and age; and \(\epsilon_{t,x}\) are independent identical distributed standard normal error terms. Equation \eqref{eq:HU_2} has a similar structure to the LC method, except that it incorporates more than one principal component and score. It describes the dynamics of \(s_t(x)\) over time. The \(\beta_{t,j}\) for every \(j\) and \(\mu(x)\) are sets of age-specific constants, and \(\phi_j(x)\) for every \(j\) is a time-varying index. The \(e_t(x)\) is the model error, which is the residual function with mean zero and no serial correlation. The \(\mu(x)\) is estimated by the average of \(s_t(x)\) across years, which measures the general shape of the observed data \(y_t(x)\). The estimates of \(\phi_j(x)\) and \(\beta_{t,j}\) are the \(j\)th principal component and scores of \(s_t(x)-\mu(x)\) respectively. In other words, we take singular value decomposition of \(s_t(x)-\mu(x)\), then \eqref{eq:HU_2} now becomes
\[
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_J(x)\beta_{t,J}
\]
We chose \(J\) in a way that we believe the majority of information for \(s_t(x)-\mu(x)\) has been captured by the first \(J\) components.

In terms of forecasting, by conditioning on the observed data \(\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}\) and the set of functional principal components \(\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}\), the \(h\)-step-ahead forecast of log rate \(y_{n+h}(x)\) can be acquired by:
\begin{equation}
  \widehat{y}_{n+h\mid n}(x)=\widehat{s}_{n+h\mid n}(x) = \text{E}[\widehat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \widehat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\widehat{\beta}_{n+h\mid n,j}
\end{equation}
where \(\widehat{\beta}_{n+h\mid n,j}\) denotes the \(h\)-step-ahead forecast of \(\beta_{n+h\mid n,j}\). Because of the time-varying property of \(\left\{\beta_{t,j}\right\}\), it controls the dynamics of the process. For every component series \(\left\{\beta_{t,j}\right\}\), we fit them with an ARIMA\((p_j,d_j,q_j)\) process, where \(p_j,d_j,q_j\) are determined by minimizing the Akaike Information Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process is given as follows:
\[
  (1-\phi_1B-\dots-\phi_{p_j}B^{p_j})(1-B)^{d_j}\beta_{t,j}=c+(1+\theta_1B+\dots+\theta_{q_j}B^{q_j})e_t, \quad ~\forall j=1,2,\dots,J,
\]
where the parameters are estimated using maximum likelihood estimation.

@@

We use the same notation and similar procedures to construct data as in \textcite{HB08}, because the fundamental model for forecasting population for this study is the \textcite{HU07} method. The data used to construct population age structure included: age- and gender-specific birth and death numbers of each calendar year, age- and gender-specific population members on 1~January of each year, and age- and gender-specific exposures to risk (i.e., population of age \(x\) at 30 June) for each year. The notations are as follows:
\begin{align*}
  B_t(x) &= \text{Births in calendar year $t$ to females of age $x$}\\
  D_t(x) &= \text{Deaths in calendar year $t$ of persons of age $x$}\\
  P_t(x) &= \text{Population of age $x$ at 1 January of year $t$}\\
  E_t(x) &= \text{Population of age $x$ exposed to risk at 30 June of year $t$}
\end{align*}
where \(x=0,1,2,\dots,p-1,p^+\) denotes age with \(p^+\) the open-ended upper age group, set to 100\(^+\) in this study. The years is denoted by \(t=1,2,\dots,n\), and superscripts \(M\) and \(F\) denote male and female respectively.

We can obtain historical mortality rates \(m_t(x)\) (also called central death rates) and fertility rates \(f_t(x)\) as follows:
\begin{align*}
  m_t(x) & = \frac{D_t(x)}{E_t(x)}=\text{age-sex-specific central death rates in calendar year t}\\
  f_t(x) & = \frac{B_t(x)}{E^F_t(x)}=\text{age-specific fertility rates in calendar year t}.
\end{align*}

In addition, the net migration\footnote{Net migration is the difference between immigration and emigration in a certain area during a specified time frame.} (denoted \(G\)) is estimated using demographic growth-balance equations, which summarize the relationship between population change and three factors: net migration, births and deaths. The equations are expressed as:
\begin{align*}
  G_t(x,x+1)     & =P_{t+1}(x+1)-P_t(x)+D_t(x,x+1)\qquad\text{for $x=0,1,2,\dots,p-2$},\\
  G_t(p-1^+,p^+) & =P_{t+1}(p^+)-P_t(p^+)-P_t(p-1)+D_t(p-1^+,p^+),\\
  G_t(B,0)       & =P_{t+1}(0)-B_t+D_t(B,0),
\end{align*}
where \(G_t(x,x+1)\) refers to net migration in calendar year \(t\) of persons aged \(x\) at beginning of year \(t\), \(G_t(p-1^+,p^+)\) refers to migration in calendar year \(t\) of persons aged \(p-1\) and older at beginning of year \(t\), and \(G_t(B,0)\) refers to deaths in calendar year \(t\) of births during year \(t\); and similarly for deaths, \(D_t(x,x+1)\), \(D_t(p-1^+,p^+)\), and \(D_t(B,0)\). The deaths are estimated using the standard life table approach of population projection. It should be noted that the estimated net migration includes errors in data recording.
